Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-01-26 paper/paper.tex
related work over

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{conjecture}{Conjecture}
\ifdef{\dpspecial}{
%% tmp by Dirk
\newcommand{\HOLtm}[1]{\texttt{#1}}
\newcommand{\HOLty}[1]{\texttt{#1}}
\newcommand{\HOLthm}[2][]{\texttt{#2}}
\newcommand{\n}{\\n}
}{}
\NewEnviron{holthmenv}{\[\begin{array}[t]{l}\BODY\end{array}\]\ignorespacesafterend}
\newcommand{\TODO}[1]{{\bf TODO:} #1}

@@

%Although full static verification is the most desirable answer to prove correctness, it is not practically possible here due to inaccessibility of source code. But even if the source were available it would not be feasible to carry such a task  because verifying every single implementation for any arbitrary STV algorithm is a Herculean significantly resource consuming agenda. Besides, any small change in implementation of an algorithm may render the whole  static verification of the implementation useless to the point that the process need to be redone from the scratch.   
We are motivated to engineer a trustworthy practical  framework to answer the following question. How can one verify that any execution of any implementation $\mathcal{P}$ whose source code is secret of an arbitrary  known STV algorithm $\mathcal{A}$ correctly computes the end result according to instructions in $\mathcal{A}$?

To begin with addressing the problem, assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ standing as a proof of correctness of the execution.
%of the election for an input $x$ consisting of election votes recorded. Given the situation, what is also sensible to ask and easy to be produce, is to make $\mathcal{P}$  also output an evidence $\omega$ of correctness of $y$. 
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of the (source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

%New questions emerge from the above problem reformulation.  
An evidence as such must be enough informative to offer transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as instructions of the counting scheme dictate so that winners truly reflect will of the voters. 
%Note that evidences are produced by various instances of computation with many STV algorithms which differ from each other in their exact counting mechanism.  
Therefore one wonders what information is enough 
%for satisfying transparency and independent verifiability of tallying 
to serve our end and what data structure for recording the information in evidence we should choose.

@@

\subsection{Characterisation of Evidence}
%%%Through thorough analysis of various STV schemes, we come to identify a common underlying data and algorithmic structure existing in the family. We abstract this structure as a finite state machine and call it the \emph{generic STV}\cite{}. Upon each execution on a given input $x$, the machine outputs a value $y$ and an \emph{evidence} $\omega$ which is the trace of all states that the machine visits from an initial machine state where $x$ is registered to the final state where $y$ is obtained.   
%%%%%Concrete STV algorithms are therefore realised as instantiations into the machine. Each of the instantiations  also produce an evidence for any instance of computation carried out. 
%%%%%%%Our objective here is to independently scrutinise process of counting under any STV schemes similar to physical scrutineers who observe tallying to attest its authenticity.  Having an evidence of computation at hand, we are then able to validate or reject the instance of computation carried out by checking if all of the transitions taken to terminate at the final state where the winners are declared are correct.  
A thorough analysis of different STV algorithms reveals existence of an underlying abstract data type common between STV algorithms. Each STV scheme operates on some ballots, each of which is a pair consisting of a preference list and a fractional value, recorded to be counted and has a quota which is the least amount of votes needed to attract in order to be elected, some candidates competing in the election, some vacancies for which candidates compete, and more importantly \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from the beginning to the end and to be able to independently verify its correctness provided access to the information is granted.

There are two kinds of discrete states of computation. Some of them are terminal states where tallying reaches to a halting state by finding the winners of the election. The other set of states consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain a terminal result. Each non-final state encloses seven pieces of data: 
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the votes that each candidate have attracted, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones whose surplus votes must be transferred,
 \item\textbf{continuing candidates} hopefully running in the election. 
 \end{itemize}
%Data recorded in the final and non-final states as specified above   record all of the information needed and sufficient to know for independently validating a counting instance.

The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$ . Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By a piece of evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
%\begin{definition}[Evidence]
%An evidence $\hat{\omega}$ for an instance of computation with an algorithm $\mathcal{A}$ is a trace of data whose abstract representation matches with some formal evidence. 
%\end{definition}
%Definition asserts the abstract data type whose values mathematically represent a state of computation. A sequence of these mathematical states formally represent an evidence. In other words, an evidence is a trace of data whose abstract characterisation is a member of the mathematical construct given in Definition. 
%\begin{definition}
%Assume $\mathit{Ballot}$, $\mathit{Tally}$, $\mathit{Pile}$ are respectively sets of ballots, tallies, and pile of ballots. Also assume $\mathcal{C}$ is the set of all of competing candidates in  an election. Then the set of final states equals to the set of all subsets of $C$, namely $2^{C}$, and the set of non-final states are 
%\begin{small}
%\begin{center}
%$\mathit{Ballot}\times\mathit{Tally}\times\mathit{Pile}\times (2^{C}\times 2^{C})\times 2^{C}\times 2^{C}$
%\end{center}
%\end{small}
%where the fourth component represents backlog of elected and removed and the last two ones are respectively for elected and continuing candidates. 
%\end{definition}
\section{The Generic STV Finite State Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already found out an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics for manipulating data to validate evidence.   On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand it must facilitate us with designing and developing the framework modularly while offering automation of synthesising process of verifiers.
%in such a way that we achieve a desirable degree of modularity and automation of  the process of synthesising different verifiers.
%evidence verifiers for various STV algorithms.

@@

In a secondary examination of the STV family, we also identify a universal algorithmic mechanism for performing computation on evidence. This generic method comprises seven general actions~(Figure~\ref{GenSteps}) corresponding to actions that tally officers take when counting votes.
\begin{figure} 
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
\caption{Generic STV Counting Steps}
\label{GenSteps}
\end{figure}
Each of the actions is constituted of  universal pre-conditions on when and how the action comes into effect, and ubiquitous post-conditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying.

@@

The underlying omnipresent abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we name \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.
% \begin{remark} 
Note that the states of computation recorded in all possible  evidences now become identical with states of the generic STV. 
%\end{remark}

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$. 
%where $S_t$ is the sanity check condition  for  transition $t \in \mathcal{T}$.  An \emph{instance} of $M$  is a tuple $I = \langle \mathcal{T}, (R_t)_{t \in \mathcal{T}} \rangle$, where for each $t \in \mathcal{T}$, $R_t \subseteq \mathcal{S} \times \mathcal{S}$ is a rule that satisfies the sanity check condition $S_t$. 
\end{definition}

\subsection{Instantiations of the Machine with Specific STV schemes}

There are nonetheless variations in the precise counting method of different STV algorithms as well. For example in the STV used for electing the upper house representatives of the Victoria state of Australia, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia does not have such pre-condition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in post-conditions. For example transfer-elect of the STV of ACT requires distributing only the last parcel of surplus votes received by an elected candidate. But the scheme of the Victoria state specifies distributing all surplus votes of an elected candidate.

@@

The operational semantics of each instantiation allows us performing operations such validation on evidences. Definition~\ref{verifier} lays down what verifying an evidence amount to. Informally speaking, to check if an evidence is valid one simply needs to inspect if transitions  between each two consecutive states of computation appearing in the evidence occur by a legitimate application of a counting action of the scheme used.

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation with $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots, |\Omega|~-1\}.$ \\
\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
\\\hspace*{0.4cm}$\Rightarrow \exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash\mathcal{S}'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(S)$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition}

We shall next provide an example of an evidence to concretely illustrate them and exemplify how counting with an STV scheme works.

@@

%%%%%%%%The process of distributing surplus votes from elected candidates and excluding the candidate with the fewest votes continues until all vacancies are filled.
%\end{description}
 \subsection{An Example of Evidence}
\label{subsec:InstEv}
Figure~\ref{EvInst} depicts evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

First preferences for each candidate are computed, and ballots counted in favour of particular candidates are placed onto that candidate’s pile. Here, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, and $b$ receives $b_{4}$, and $c$ receives $b_{5}$ . Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. $b$ and $c$ each attracts one vote. However  neither of them reaches to the quota. Therefore the weakest one $b$ is eliminated. As the number of elected and continuing candidates does not exceed the initial vacancies, the election terminates by declaring $a$ and $c$ winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

  % STATE STVONE
 \node[state,
  below of=Aux,
  %yshift = -2cm,
  node distance=1.3cm,
  anchor=center] (proof) 
 {%
 \begin{tabular}{l}
  \textbf{Proofs}
 \end{tabular}
 };

@@

  % STATE STVTWO
 \node[state,
  right of= proof,
  xshift = 1.3cm,
  %node distance=1.5cm,
  anchor=center] (STV2) 
 {%
 \begin{tabular}{l}
  \textbf{STV$_{2}$}
 \end{tabular}
 };
   % STATE EXETWO
 \node[state,
  below of= STV2,
  xshift = 1cm,
  node distance=1.3cm,
  anchor=center] (V) 
 {%
 \begin{tabular}{l}
  \textbf{Verifier}
 \end{tabular}
 };

   \node[state,
  below of= STV1,
  xshift = - 1cm,
  node distance=1.3cm,
  anchor=center] (T) 
 {%
 \begin{tabular}{l}
  \textbf{Tran}
 \end{tabular}
 };

 \node[state,
  below of = proof,
   %xshift = 3cm,
   node distance = 1.1cm,
  anchor=center
  %text width=2.5cm
  ] (Deep) 
 {%
 \begin{tabular}{l}
  \textbf{DeepSpec}
%  \parbox{1.7cm}{}
 \end{tabular}
 };

@@

\end{tikzpicture}
\caption{Architecture of the Framework}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise $x86\_64$ executable  verifier $\mathcal{V}_{i}$.

@@

In the same manner as elaborated for transfer-elect, we obtain specification, computational implementation for other machine transitions as well. Then the specification (implementation) of the machine semantics is comprised of the collection of specification (resp. implementation) of the individual transitions. We formally prove that for each transition, the implementation corresponds with its specification.
%Drawing on equivalences  established for the components, we demonstrate the correspondence  between the specification and implementation of the transitions semantics.

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{conjecture}{Desideratum}
\ifdef{\dpspecial}{
%% tmp by Dirk
\newcommand{\HOLtm}[1]{\texttt{#1}}
\newcommand{\HOLty}[1]{\texttt{#1}}
\newcommand{\HOLthm}[2][]{\texttt{#2}}
\newcommand{\n}{\\n}
}{}
\NewEnviron{holthmenv}{\[\begin{array}[t]{l}\BODY\end{array}\]\ignorespacesafterend}
\newcommand{\TODO}[1]{{\bf TODO:} #1}

@@

%Although full static verification is the most desirable answer to prove correctness, it is not practically possible here due to inaccessibility of source code. But even if the source were available it would not be feasible to carry such a task  because verifying every single implementation for any arbitrary STV algorithm is a Herculean significantly resource consuming agenda. Besides, any small change in implementation of an algorithm may render the whole  static verification of the implementation useless to the point that the process need to be redone from the scratch.   
We are motivated to 
%engineer a trustworthy practical  framework to answer the 
tackle the following problem. How can one verify that any execution of any implementation $\mathcal{P}$ whose source code is secret of an arbitrary  known STV algorithm $\mathcal{A}$ correctly computes the end result according to  $\mathcal{A}$?

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.
%of the election for an input $x$ consisting of election votes recorded. Given the situation, what is also sensible to ask and easy to be produce, is to make $\mathcal{P}$  also output an evidence $\omega$ of correctness of $y$. 
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of the (source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

%New questions emerge from the above problem reformulation.  
The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. 
%Note that evidences are produced by various instances of computation with many STV algorithms which differ from each other in their exact counting mechanism.  
Therefore the questions becomes what information is enough 
%for satisfying transparency and independent verifiability of tallying 
for establishing transparency and verifiability of tallying  and what data structure for recording the information in evidence we should choose.

@@

\subsection{Characterisation of Evidence}
%%%Through thorough analysis of various STV schemes, we come to identify a common underlying data and algorithmic structure existing in the family. We abstract this structure as a finite state machine and call it the \emph{generic STV}\cite{}. Upon each execution on a given input $x$, the machine outputs a value $y$ and an \emph{evidence} $\omega$ which is the trace of all states that the machine visits from an initial machine state where $x$ is registered to the final state where $y$ is obtained.   
%%%%%Concrete STV algorithms are therefore realised as instantiations into the machine. Each of the instantiations  also produce an evidence for any instance of computation carried out. 
%%%%%%%Our objective here is to independently scrutinise process of counting under any STV schemes similar to physical scrutineers who observe tallying to attest its authenticity.  Having an evidence of computation at hand, we are then able to validate or reject the instance of computation carried out by checking if all of the transitions taken to terminate at the final state where the winners are declared are correct.  
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\begin{itemize}
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to attract in order to be elected, 
\item some candidates competing in the election, 
\item some number of vacancies for which candidates compete,
\item \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from the beginning to the end and to be able to independently verify its correctness provided access to the information is granted. 
\end{itemize}
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
 \end{itemize}
%Data recorded in the final and non-final states as specified above   record all of the information needed and sufficient to know for independently validating a counting instance.

The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$ . Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
%\begin{definition}[Evidence]
%An evidence $\hat{\omega}$ for an instance of computation with an algorithm $\mathcal{A}$ is a trace of data whose abstract representation matches with some formal evidence. 
%\end{definition}
%Definition asserts the abstract data type whose values mathematically represent a state of computation. A sequence of these mathematical states formally represent an evidence. In other words, an evidence is a trace of data whose abstract characterisation is a member of the mathematical construct given in Definition. 
%\begin{definition}
%Assume $\mathit{Ballot}$, $\mathit{Tally}$, $\mathit{Pile}$ are respectively sets of ballots, tallies, and pile of ballots. Also assume $\mathcal{C}$ is the set of all of competing candidates in  an election. Then the set of final states equals to the set of all subsets of $C$, namely $2^{C}$, and the set of non-final states are 
%\begin{small}
%\begin{center}
%$\mathit{Ballot}\times\mathit{Tally}\times\mathit{Pile}\times (2^{C}\times 2^{C})\times 2^{C}\times 2^{C}$
%\end{center}
%\end{small}
%where the fourth component represents backlog of elected and removed and the last two ones are respectively for elected and continuing candidates. 
%\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics for manipulating data to validate evidence.   On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand it must facilitate designing and developing the framework modularly while offering automation of synthesising process of verifiers.
%in such a way that we achieve a desirable degree of modularity and automation of  the process of synthesising different verifiers.
%evidence verifiers for various STV algorithms.

@@

In a secondary examination of the STV family, we also identify a general algorithmic mechanism for performing computation on evidence, which realises the \emph{commonalities} among the family members. This generic method comprises seven general actions~(Figure~\ref{GenSteps}) corresponding to actions that tally officers take when counting votes.
\begin{figure} 
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
\caption{Generic STV Counting Steps}
\label{GenSteps}
\end{figure}
Each action is constituted of  pre-conditions on when and how the action comes into effect, and post-conditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying, and they are common between all flavours of STV.

@@

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.
% \begin{remark} 
Note that the states of computation recorded in all possible  evidence now become identical with machine states.
%states of the generic STV. 
%\end{remark}

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$. 
%where $S_t$ is the sanity check condition  for  transition $t \in \mathcal{T}$.  An \emph{instance} of $M$  is a tuple $I = \langle \mathcal{T}, (R_t)_{t \in \mathcal{T}} \rangle$, where for each $t \in \mathcal{T}$, $R_t \subseteq \mathcal{S} \times \mathcal{S}$ is a rule that satisfies the sanity check condition $S_t$. 
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  
 For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV), distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such pre-condition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in post-conditions. For example transfer-elect of the ACT STV used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

The operational semantics of each instantiation allows  to perform operations such as validation on evidence. Definition~\ref{verifier} lays down what verifying an evidence amount to. Informally speaking, to check if  evidence is valid one simply needs to inspect if transitions  between each two consecutive states of computation appearing in the evidence occur by a legitimate application of a counting action of the scheme used.

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation with $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash\mathcal{S}'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(S)$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition}

We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work.

@@

%%%%%%%%The process of distributing surplus votes from elected candidates and excluding the candidate with the fewest votes continues until all vacancies are filled.
%\end{description}
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
Figure~\ref{EvInst} depicts evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

First preferences for each candidate are computed, and ballots counted in favour of particular candidates are placed onto that candidate’s pile. Here, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, and $b$ receives $b_{4}$, and $c$ receives $b_{5}$ . Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attracts one vote. However  neither of them reaches to the quota. Therefore the weakest one $b$ is eliminated. As the number of elected and continuing candidates does not exceed the initial vacancies, the election terminates by declaring $a$ and $c$ winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

  % STATE STVONE
 \node[state,
  below of=Aux,
  %yshift = -2cm,
  node distance=1.1cm,
  anchor=center] (proof) 
 {%
 \begin{tabular}{l}
  \textbf{Proofs}
 \end{tabular}
 };

@@

  % STATE STVTWO
 \node[state,
  right of= proof,
  xshift = 1.2cm,
  %node distance=1.5cm,
  anchor=center] (STV2) 
 {%
 \begin{tabular}{l}
  \textbf{STV$_{2}$}
 \end{tabular}
 };
   % STATE EXETWO
 \node[state,
  below of= STV2,
  xshift = 1cm,
  node distance=1.2cm,
  anchor=center] (V) 
 {%
 \begin{tabular}{l}
  \textbf{Verifier}
 \end{tabular}
 };

   \node[state,
  below of= STV1,
  xshift = - 1cm,
  node distance=1.2cm,
  anchor=center] (T) 
 {%
 \begin{tabular}{l}
  \textbf{Tran}
 \end{tabular}
 };

 \node[state,
  below of = proof,
   %xshift = 3cm,
   node distance = 1cm,
  anchor=center
  %text width=2.5cm
  ] (Deep) 
 {%
 \begin{tabular}{l}
  \textbf{DeepSpec}
%  \parbox{1.7cm}{}
 \end{tabular}
 };

@@

\end{tikzpicture}
\caption{Architecture of the Framework, note that direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.

@@

Similarly 
%as elaborated for transfer-elect, 
we obtain specification and computational implementation for other machine transitions as well. Then the specification (implementation) of the machine semantics is comprised of the collection of specification (resp. implementation) of the individual transitions. We formally prove 
%formally prove that for 
each transition implementation  matches with its specification.
%Drawing on equivalences  established for the components, we demonstrate the correspondence  between the specification and implementation of the transitions semantics.
